Revisiting `Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation

### 2023-05-24 analysis/quarto/paper.qmd
incorporate another set of comments and calculate overall page times for cost estimation

---
title: Revisiting 'Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  I revisit the dynamic public goods game by Gächter et al. (2017) developed to study cooperation under dynamic interdependencies. Collecting data from both a convenient (students) and an inconvenient (general population) sample, I not only reproduce some of the author's original observations but also test their novel game's generalizability. Appending a charitable dictator game, I find no correlations between behavior in the charitable context and the dynamic game. This applies to students and the general population sample alike. Because the study of inexperiend general population samples raises methodological challenges, such as fatigue and dropouts, this research approaches them. Doing so, I provide simple solutions to run relieable interactive experiments online.

@@

Public good (or public bad) games---although often intended to inform climate policies (e.g. @MilinskiEtAl2006, @TavoniEtAl2011, @Hauser2014, @BrickEtAl2015, @GomezEtAl2018, @CalzolariEtAl2018, @CookEtAl2019)---miss these temporal interdependencies simply because participants have the same set of actions in each period. Accordingly, participants' actions in a given period do not affect their action space in subsequent periods. To see the lack of realism, consider carbon dioxide emissions, where the current stock will last for well over a millennium [@Inman2008; @CalzolariEtAl2018]. Playing with fresh endowments in each period is as if one could just undo carbon dioxide emissions at no cost.

A game designed by Gächter, Mengel, Tsakas & Vostroknutov [-@GMTV2017 hereafter, GMTV] as well as Stefan Große (2011, unpublished) shows that it is fairly simple to remedy that limitation. They incorporated interdependencies into a _dynamic_ public goods game by defining endowments as the income of previous periods. To put it differently, participants' actions in a given period affect their number of actions in subsequent periods: the more (less) they earn now, the more (less) they can contribute in the next period. Importantly, this modification qualitatively yields the same rational predictions as the static game (i.e. free-riding and the under-provision of the public good). It is thus, equally well suited to study dilemma situations.

@@

@AGM2018 conducted static public goods games in the lab and on MTurk to draw lessons from online experimentation. This study extends this literature [see @GoodmanPaolacci2017; also  @AmirEtAl2012] by focusing on an _inconvenient_ sample (i.e. one that is completely inexperienced sample that has not been exposed to interactive experiments before) playing a computationally more complex game. This required me to design a robust (and thus, more complex) software to minimize attrition. I collected paradata to assess the desired fluency and feasibility of the experiment. This research reports on the robust design that made the dynamic game feasible.

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations were established to tie the researchers' hands when doing confirmatory research--reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b). Today, researchers can not only choose where to pre-register but also decide about the level of detail they provide. More severely, nothing stops researchers from pre-registering multiple hypotheses opposing each other (see, e.g., @SimmonsEtAl2021 and @PhamEtAl2021 for a discussion).

@@

In fact, the experiment reported in this article qualifies for the _'pre-registered'_ label at a first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run when I designed the experiment on [GitHub](https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd).^[https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd to run the code, you need to executed the .Rmd files in this repository in the order that is indicated by its file names.]

@@

Accordingly, I  changed many of the analysis scripts.^[The source code of this document contains the analysis and can be found [here](https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd): https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd] However, because the originally planned analysis code is archived and reproducible, these changes are transparent. Importantly, I followed a literate programming approach [@Knuth_1984; @AkhtarYe_2023]. Hence, all documents needed to analyze the data and to write the report stem from the same source. This establishes consistence between the commands one tells the statistical software to do and the explanation one tells human beings one told the statistical software to do. As such, (a), (b) and the transition from (a) to (b) are not only transparent but also comprehensible.

@@

**Result 3.** _There is no significant correlation between average contributions in the abstract public goods game and contributions to the real public good of climate change mitigation._

>>>>

---
title: Revisiting 'Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  I revisit the dynamic public goods game developed by Gächter et al. (2017) to study cooperation under dynamic interdependencies. Collecting data from both a convenient (students) and an inconvenient (general population) sample, I not only reproduce some of the author's original observations but also test their novel game's generalizability. Appending a charitable dictator game, I find no correlations between behavior in the charitable context and the dynamic game. This applies to students and the general population sample alike. Because the study of inexperiend general population samples raises methodological challenges, such as fatigue and dropouts, this research approaches them. Doing so, I provide simple solutions to run relieable interactive experiments online.

@@

Public good (or public bad) games---although often intended to inform climate policies [e.g. @MilinskiEtAl2006; @TavoniEtAl2011; @Hauser2014; @BrickEtAl2015; @GomezEtAl2018; @CalzolariEtAl2018; @CookEtAl2019]---miss these temporal interdependencies simply because participants have the same set of actions in each period. Accordingly, participants' actions in a given period do not affect their action space in subsequent periods. To see the lack of realism, consider carbon dioxide emissions, where the current stock will last for well over a millennium [@Inman2008; @CalzolariEtAl2018]. Playing with fresh endowments in each period is as if one could just undo carbon dioxide emissions at no cost.

A game designed by Gächter, Mengel, Tsakas & Vostroknutov [-@GMTV2017 hereafter, GMTV] as well as Stefan Große (2011, unpublished) shows that it is fairly simple to remedy that limitation. They incorporated interdependencies into a _dynamic_ public goods game (dPGG) by defining endowments as the income of previous periods. To put it differently, participants' actions in a given period affect their number of actions in subsequent periods: the more (less) they earn now, the more (less) they can contribute in the next period. Importantly, this modification qualitatively yields the same rational predictions as the static game (i.e. free-riding and the under-provision of the public good). It is thus, equally well suited to study dilemma situations.

@@

@AGM2018 conducted static public goods games in the lab and on MTurk to draw lessons from online experimentation. This study extends this literature [see, e.g., @SnowbergYariv_2021; @GuptaRigottiWilson_2021; @GoodmanPaolacci2017; @AmirEtAl2012] by focusing on an _inconvenient_ sample (i.e. one that is completely inexperienced sample that has not been exposed to interactive experiments before) playing a computationally more complex game. This required me to design a robust (and thus, more complex) software to minimize attrition. I collected paradata to assess the desired fluency and feasibility of the experiment. This research reports on the robust design that made the dynamic game feasible.

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations were established to tie the researchers' hands when doing confirmatory research--reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b). Today, researchers can not only choose where to pre-register but also decide about the level of detail they provide. More severely, nothing stops researchers from pre-registering multiple hypotheses opposing each other [see, e.g., @SimmonsEtAl2021; and @PhamEtAl2021 for a discussion].

@@

In fact, the experiment reported in this article qualifies for the _'pre-registered'_ label at a first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run when I designed the experiment on [GitHub](https://shorturl.at/bAGKY).^[https://shorturl.at/bAGKY to run the code, you need to execute the .Rmd files in this repository in the order that is indicated by file names.]

@@

Accordingly, I edited many of the analyses scripts.^[The source code of this document contains the analysis and can be found [here](https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd): https://shorturl.at/ilpAG] However, because the originally planned analysis code is archived and reproducible, these changes are transparent. Importantly, I followed a literate programming approach [@Knuth_1984; @AkhtarYe_2023]. Hence, all documents needed to analyze the data and to write the report stem from the same source. This establishes consistence between the commands one tells the statistical software to do and the explanation one tells human beings one told the statistical software to do. As such, (a), (b) and the transition from (a) to (b) are not only transparent but also comprehensible.

@@

**Result 3.** _There is no significant correlation between average contributions in the abstract public goods game and contributions to the real public good of climate change mitigation---for none of the samples._

### 2023-05-24 analysis/quarto/paper.tex
incorporate another set of comments and calculate overall page times for cost estimation

\begin{abstract}
I revisit the dynamic public goods game by Gächter et al.~(2017)
developed to study cooperation under dynamic interdependencies.
Collecting data from both a convenient (students) and an inconvenient
(general population) sample, I not only reproduce some of the author's
original observations but also test their novel game's generalizability.
Appending a charitable dictator game, I find no correlations between
behavior in the charitable context and the dynamic game. This applies to
students and the general population sample alike. Because the study of
inexperiend general population samples raises methodological challenges,
such as fatigue and dropouts, this research approaches them. Doing so, I
provide simple solutions to run relieable interactive experiments
online.

@@

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public goods game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, interior hidden, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

Public good (or public bad) games---although often intended to inform
climate policies (e.g. \citet{MilinskiEtAl2006}, \citet{TavoniEtAl2011},
\citet{Hauser2014}, \citet{BrickEtAl2015}, \citet{GomezEtAl2018},
\citet{CalzolariEtAl2018}, \citet{CookEtAl2019})---miss these temporal
interdependencies simply because participants have the same set of
actions in each period. Accordingly, participants' actions in a given
period do not affect their action space in subsequent periods. To see
the lack of realism, consider carbon dioxide emissions, where the
current stock will last for well over a millennium
\citep{Inman2008, CalzolariEtAl2018}. Playing with fresh endowments in
each period is as if one could just undo carbon dioxide emissions at no
cost.

A game designed by Gächter, Mengel, Tsakas \& Vostroknutov \citeyearpar[
hereafter, GMTV]{GMTV2017} as well as Stefan Große (2011, unpublished)
shows that it is fairly simple to remedy that limitation. They
incorporated interdependencies into a \emph{dynamic} public goods game
by defining endowments as the income of previous periods. To put it
differently, participants' actions in a given period affect their number
of actions in subsequent periods: the more (less) they earn now, the
more (less) they can contribute in the next period. Importantly, this
modification qualitatively yields the same rational predictions as the
static game (i.e.~free-riding and the under-provision of the public
good). It is thus, equally well suited to study dilemma situations.

@@

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw lessons from online experimentation. This study extends
this literature
\citetext{\citealp[see][]{GoodmanPaolacci2017}; \citealp[also][]{AmirEtAl2012}}
by focusing on an \emph{inconvenient} sample (i.e.~one that is
completely inexperienced sample that has not been exposed to interactive
experiments before) playing a computationally more complex game. This
required me to design a robust (and thus, more complex) software to
minimize attrition. I collected paradata to assess the desired fluency
and feasibility of the experiment. This research reports on the robust
design that made the dynamic game feasible.

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations were established to tie the
researchers' hands when doing confirmatory research--reducing observable
discrepancies between Daryl Bem's above-mentioned (a) and (b). Today,
researchers can not only choose where to pre-register but also decide
about the level of detail they provide. More severely, nothing stops
researchers from pre-registering multiple hypotheses opposing each other
(see, e.g., \citet{SimmonsEtAl2021} and \citet{PhamEtAl2021} for a
discussion).

@@

In fact, the experiment reported in this article qualifies for the
\emph{`pre-registered'} label at a first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run when I designed the experiment on
\href{https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd}{GitHub}.\footnote{https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd
  to run the code, you need to executed the .Rmd files in this
  repository in the order that is indicated by its file names.}

@@

Accordingly, I changed many of the analysis scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}
However, because the originally planned analysis code is archived and
reproducible, these changes are transparent. Importantly, I followed a
literate programming approach \citep{Knuth_1984, AkhtarYe_2023}. Hence,
all documents needed to analyze the data and to write the report stem
from the same source. This establishes consistence between the commands
one tells the statistical software to do and the explanation one tells
human beings one told the statistical software to do. As such, (a), (b)
and the transition from (a) to (b) are not only transparent but also
comprehensible.

@@

\textbf{Result 3.} \emph{There is no significant correlation between
average contributions in the abstract public goods game and
contributions to the real public good of climate change mitigation.}

>>>>

\begin{abstract}
I revisit the dynamic public goods game developed by Gächter et
al.~(2017) to study cooperation under dynamic interdependencies.
Collecting data from both a convenient (students) and an inconvenient
(general population) sample, I not only reproduce some of the author's
original observations but also test their novel game's generalizability.
Appending a charitable dictator game, I find no correlations between
behavior in the charitable context and the dynamic game. This applies to
students and the general population sample alike. Because the study of
inexperiend general population samples raises methodological challenges,
such as fatigue and dropouts, this research approaches them. Doing so, I
provide simple solutions to run relieable interactive experiments
online.

@@

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public goods game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, interior hidden, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

Public good (or public bad) games---although often intended to inform
climate policies
\citep[e.g.][]{MilinskiEtAl2006, TavoniEtAl2011, Hauser2014, BrickEtAl2015, GomezEtAl2018, CalzolariEtAl2018, CookEtAl2019}---miss
these temporal interdependencies simply because participants have the
same set of actions in each period. Accordingly, participants' actions
in a given period do not affect their action space in subsequent
periods. To see the lack of realism, consider carbon dioxide emissions,
where the current stock will last for well over a millennium
\citep{Inman2008, CalzolariEtAl2018}. Playing with fresh endowments in
each period is as if one could just undo carbon dioxide emissions at no
cost.

A game designed by Gächter, Mengel, Tsakas \& Vostroknutov \citeyearpar[
hereafter, GMTV]{GMTV2017} as well as Stefan Große (2011, unpublished)
shows that it is fairly simple to remedy that limitation. They
incorporated interdependencies into a \emph{dynamic} public goods game
(dPGG) by defining endowments as the income of previous periods. To put
it differently, participants' actions in a given period affect their
number of actions in subsequent periods: the more (less) they earn now,
the more (less) they can contribute in the next period. Importantly,
this modification qualitatively yields the same rational predictions as
the static game (i.e.~free-riding and the under-provision of the public
good). It is thus, equally well suited to study dilemma situations.

@@

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw lessons from online experimentation. This study extends
this literature \citep[see,
e.g.,][]{SnowbergYariv_2021, GuptaRigottiWilson_2021, GoodmanPaolacci2017, AmirEtAl2012}
by focusing on an \emph{inconvenient} sample (i.e.~one that is
completely inexperienced sample that has not been exposed to interactive
experiments before) playing a computationally more complex game. This
required me to design a robust (and thus, more complex) software to
minimize attrition. I collected paradata to assess the desired fluency
and feasibility of the experiment. This research reports on the robust
design that made the dynamic game feasible.

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations were established to tie the
researchers' hands when doing confirmatory research--reducing observable
discrepancies between Daryl Bem's above-mentioned (a) and (b). Today,
researchers can not only choose where to pre-register but also decide
about the level of detail they provide. More severely, nothing stops
researchers from pre-registering multiple hypotheses opposing each other
\citetext{\citealp[see, e.g.,][]{SimmonsEtAl2021}; \citealp[and][ for a
discussion]{PhamEtAl2021}}.

@@

In fact, the experiment reported in this article qualifies for the
\emph{`pre-registered'} label at a first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run when I designed the experiment on
\href{https://shorturl.at/bAGKY}{GitHub}.\footnote{https://shorturl.at/bAGKY
  to run the code, you need to execute the .Rmd files in this repository
  in the order that is indicated by file names.}

@@

Accordingly, I edited many of the analyses scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://shorturl.at/ilpAG} However, because the originally planned
analysis code is archived and reproducible, these changes are
transparent. Importantly, I followed a literate programming approach
\citep{Knuth_1984, AkhtarYe_2023}. Hence, all documents needed to
analyze the data and to write the report stem from the same source. This
establishes consistence between the commands one tells the statistical
software to do and the explanation one tells human beings one told the
statistical software to do. As such, (a), (b) and the transition from
(a) to (b) are not only transparent but also comprehensible.

@@

\textbf{Result 3.} \emph{There is no significant correlation between
average contributions in the abstract public goods game and
contributions to the real public good of climate change mitigation---for
none of the samples.}
